---
title: "From slide measurements to a traceable report"
output:
  rmarkdown::html_vignette:
    toc: true
    css: style.css
    number_sections: false
vignette: >
  %\VignetteIndexEntry{From slide measurements to a traceable report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(bloodspottR)
```

## Scope and measurement meaning

This workflow starts with physical-slide measurements. It does not read native
whole slides, train a DNN or validate a candidate as one erythrocyte. The reusable
package preserves distinctions needed for those later stages: measured red area,
spot candidates, unresolved bulk events, explicit support and provenance.
Software validation checks contracts and arithmetic, not biological accuracy.

The examples use deterministic **synthetic** data. One record is deliberately
missing signal measurements and another has zero measured tissue support.

```{r example}
results <- bs_example()
bs_validate(results)
results$slides[, c("physical_slide_id", "organ", "tissue_area_mm2",
                   "red_area_mm2", "candidate_spots")]
```

Canonical tables identify physical slides with nonempty character strings, areas
in square millimetres and nonnegative whole-valued counts. The operational count
is candidate spots plus unresolved bulk events. Counts use numeric storage and
must remain exactly representable within the documented double-precision bound.
Existing QC/organ labels are metadata, never inferred from signal strength.

## Pool numerators and support

A group's red-positive fraction is
`100 * sum(red_area_mm2) / sum(tissue_area_mm2)`.
Its candidate density uses the same tissue denominator. Pooled rates are not
means of slide rates: a small and a large slide contribute different support.

```{r pooled}
by_organ <- bs_summarize(results, strata = "organ")
by_organ[, c("organ", "n_slides", "tissue_area_mm2", "red_area_mm2",
             "red_percent", "spots_per_mm2")]
is.na(by_organ$red_percent[by_organ$organ == "Spleen"])
```

The spleen summary is missing because one contributing measurement is missing.
It would be misleading to silently sum only its known values while retaining the
full tissue denominator. A true zero numerator on positive support is zero;
zero or unknown support gives a missing rate. Excluding incomplete or compromised
slides requires a declared analysis decision and a retained membership record.

Group jointly by tissue and the supplied quality strata:

```{r strata}
results$groups <- bs_summarize(results, strata = c("organ", "stratum"))
results$groups[, c("organ", "stratum", "n_slides", "red_percent")]
```

Only observed combinations appear. `slide_ids` retains membership. A
`qc_revision` can replace metadata when it uniquely covers every measured slide;
it cannot replace numerical measurements. Historical study-specific thresholds
and exclusion decisions are not imposed by this package.

## Import a preserved measurement table

The legacy importer accepts a canonical CSV or a saved `analysis.json` with a
`rows` table. The `p079` profile also maps `tissue_mm2` and `red_mm2`; it does not
assign controls or invent missing calibration. Source file identity is hashed.

```{r import}
source_file <- tempfile(fileext = ".csv")
utils::write.csv(results$slides, source_file, row.names = FALSE, na = "")
imported <- bs_import_legacy(source_file, profile = "canonical")
identical(imported$slides$physical_slide_id, results$slides$physical_slide_id)
bs_status(imported)
unlink(source_file)
```

The source hash is an integrity receipt, not proof of authentic acquisition or
independent biological validation. Canonical result JSON exported by the package
contains a schema-versioned `slides` table for portable exchange. The importer
accepts that format as well as the historical `rows` format.

## Compare runs on declared common support

Equal slide IDs alone do not prove equal measurement support. Here both runs
explicitly reference the same synthetic support identity, so paired changes can
be computed. Without matching support, `support = "common"` fails; the union
view preserves unmatched records with unverifiable differences missing.

```{r compare}
first <- results
second <- results
first$slides$support_id <- paste0("synthetic-support-", first$slides$physical_slide_id)
second$slides$support_id <- first$slides$support_id
second$slides$red_area_mm2[2] <- 0.25
comparison <- bs_compare(list(baseline = first, comparison = second))
comparison[, c("physical_slide_id", "support_verified", "red_area_mm2_delta")]
results$comparison <- comparison
```

## Preserve a report and tables

Each delivery uses a new destination directory. Existing deliveries are never
overwritten; the parent directory must exist. HTML text is escaped, exports keep
missing values, and a SHA-256 manifest records generated files.

```{r report}
delivery <- tempfile("bloodspottr-example-report-")
report <- bs_report(results, delivery, author = "RH",
                    background = "Synthetic demonstration; no biological inference.")
list.files(delivery)
file.exists(c(report, file.path(delivery, "SHA256SUMS")))
```

`bs_export_results(results, out)` writes tables without HTML. Set `xlsx = TRUE`
for optional **openxlsx2** workbook output. The initial renderer does not embed
raw image masks, native detail crops, training weights or individual events; the
study-specific TeX/PDF supplement is a separate artifact.

```{r optional-xlsx}
if (requireNamespace("openxlsx2", quietly = TRUE)) {
  workbook_delivery <- tempfile("bloodspottr-example-xlsx-")
  bs_export_results(results, workbook_delivery, xlsx = TRUE)
  file.exists(file.path(workbook_delivery, "results.xlsx"))
  unlink(workbook_delivery, recursive = TRUE)
}
```

## Explore with Shiny

```{r shiny, eval=FALSE}
bs_app(results)                  # Local interactive session
app <- bs_app(launch = FALSE)    # Empty application object; explicit demo button
```

The application filters by tissue, recorded QC stratum and literal slide ID.
Pooled cards and plots follow the selected rows; the source object is unchanged.
It exposes exact measurements, supplied comparisons, provenance and filtered CSV
exports. Its explicit demonstration label prevents synthetic values from being
mistaken for a loaded study. Browser uploads accept CSV/JSON, not executable
serialized objects. It does not launch a GPU backend or claim native annotation
editing or model training controls.

```{r cleanup, include=FALSE}
unlink(delivery, recursive = TRUE)
```
