---
title: "Scoped interoperability across the CTTIR suite"
output:
  rmarkdown::html_vignette:
    toc: true
    css: style.css
    number_sections: false
vignette: >
  %\VignetteIndexEntry{Scoped interoperability across the CTTIR suite}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(bloodspottR)
```

## Reuse without hidden execution

bloodspottR owns the histology workflow, calibrated burden summaries and event
semantics. Packages elsewhere in the suite own their respective image, annotation,
segmentation, spatial, cell-data or reporting facilities. A source inventory
records available package identities and exports. Presence in that inventory is not a promise that every package has been installed,
invoked or qualified end to end.

The inventory reads immediate child package metadata and literal namespace
exports. It does not load packages, execute their source, install dependencies or
start services. Directory names and declared package names can differ; duplicate
identities remain visible rather than silently overwriting one another.

```{r inventory-example}
suite <- tempfile("cttir-inventory-example-")
dir.create(suite)
dir.create(file.path(suite, "example-source"))
writeLines(c("Package: exampleTool", "Version: 0.1.0", "Title: Metadata Fixture"),
           file.path(suite, "example-source", "DESCRIPTION"))
writeLines("export(example_function)",
           file.path(suite, "example-source", "NAMESPACE"))
inventory <- bs_cttir_inventory(suite)
inventory[, c("directory", "package", "version", "exports", "status")]
inventory$status == "metadata_only"
unlink(suite, recursive = TRUE)
```

On a workstation, inventory an explicitly chosen suite root:

```{r local-suite, eval=FALSE}
inventory <- bs_cttir_inventory("/path/to/CTTIR/public", check_installed = TRUE)
inventory[inventory$duplicate_package, ]
```

An installed version may differ from source. Before a production adapter call,
record the exact version, input units, identity mapping and runtime verification.
Source inspection cannot substitute for exercising native clients and optional
backends with representative inputs.

## Calibrated event exchange with cellspecR

The implemented `bs_as_cellspec()` adapter exchanges candidate centres and image
metadata. A physical slide becomes `sample_id`; no animal or treatment assignment
is inferred. `event_id` becomes `cell_id` for schema compatibility while
`object_type = "event_candidate"` and provenance preserve its actual meaning.
It supplies neither invented cell boundaries nor intensity measurements.

```{r calibrated-events}
images <- data.frame(
  image_id = "image-001", physical_slide_id = "slide-001",
  pixel_width_um = 0.25, pixel_height_um = 0.5,
  width_px = 200L, height_px = 100L
)
events <- data.frame(
  event_id = c("candidate-001", "candidate-002"),
  image_id = "image-001", physical_slide_id = "slide-001",
  x_um = c(5, 20), y_um = c(10, 30),
  event_type = c("single_compatible", "unresolved_bulk")
)
```

Coordinates already refer to the declared image frame in micrometres. Pixel
coordinates divide each axis by its own calibration: anisotropic sampling is
not forced into a square-pixel assumption. Centres must lie inside half-open
image extents, image/slide identities must agree, and identifiers remain strings.

```{r cellspec-optional}
if (requireNamespace("cellspecR", quietly = TRUE)) {
  exchanged <- bs_as_cellspec(events, images,
                              provenance = list(source = "synthetic example"))
  class(exchanged)
}
```

This adapter does not perform spatial registration. A crop or tile first needs a
verified transform into the declared frame. Repeat acquisitions of one physical
area do not create additional biological area. Data interchange preserves these
semantics; it cannot establish geometric independence or count accuracy.

## Compatibility boundaries

Generic JSON/CSV result exports carry summaries and provenance without requiring
the image toolchain. Other scope-appropriate adapters can build on explicit
contracts: segmentation products need calibrated masks; annotation exchange
needs parent-coordinate identity and class-specific review scope; spatial
analyses need observed tissue support rather than image-canvas area.

An unavailable optional dependency should produce an actionable error. It must
not silently replace a trained model, fabricate a zero, infer missing control
labels or re-run an old study. Scope extensions belong in the package that owns
the corresponding data contract and require focused round-trip tests there.
