---
title: "Saved reviews and reproducible external workers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Saved reviews and reproducible external workers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(bloodspottR)
```

## Freeze the review pool

A review queue binds each image to its parent and SHA-256 checksum. Its default
budget is 50 images. Sampling is reproducible and preserves the caller's random
state. The caller must define the sampling pool and explain how it was obtained;
a bounded queue does not establish that examples represent all slides.

```{r queue}
work <- tempfile("review-example-")
dir.create(work)
image <- file.path(work, "field.png")
grDevices::png(image, width = 200, height = 200)
par(mar = rep(0, 4))
plot.new()
text(.5, .5, "Synthetic empty field")
grDevices::dev.off()
queue <- bs_review_plan(
  data.frame(item_id = "field-1", parent_id = "slide-1", path = image),
  budget = 50, seed = 41, out = file.path(work, "review")
)
queue$items[c("item_id", "parent_id", "selection")]
```

## Save in the annotation client, then confirm

The package does not automate a running QuPath window. Save in the client and
export a CSV with `class`, `x`, and `y` in the parent image's pixel frame before
confirmation. Crop offsets and transforms belong to the client adapter; this
ledger cannot recover them from points. Supported classes are `Spot_center`,
`Missed_spot`, `False_positive`, `Review_uncertain`, `Missed_area`,
`False_positive_area`, `Cell_nucleus`, and `Vessel_structure`.

A point identifies a location, not an area boundary. Exemplary nucleus points do
not label unmarked nuclei as absent. Vessel-wall points are not lumen masks.
Review scope is recorded separately for erythrocytes, nuclei, and vessels.
A header-only file becomes an explicit negative review only with an `absent`
assertion; it is never silently treated as a negative training example.

```{r confirm}
annotation <- file.path(work, "saved-points.csv")
utils::write.csv(data.frame(class = character(), x = numeric(), y = numeric()),
                 annotation, row.names = FALSE)
receipt <- bs_review_confirm(queue, "field-1", annotation,
  assertions = list(ery = "absent", nuclei = "unreviewed", vessel = "unreviewed"))
file.exists(file.path(queue$path, "receipts"))
```

Confirmations must follow queue order. Repeating an identical confirmation is
idempotent; changed source images, changed archived annotations, contradictory
scopes, or competing writers are rejected. Receipts are durable saved-export
records, not proof that a reader examined every object.

## External worker protocol

`bs_backend()` configures an explicitly trusted executable. `bs_train()` and
`bs_analyze()` run that executable through `processx`; they do not contain a
bundled DNN architecture or silently install Python. The worker receives two
final arguments: a JSON request path and an empty staging directory. The request
records operation, backend version, named input paths and hashes, and parameters.
A successful worker writes `result.json` with this structure:

```json
{"schema_version":1,"status":"complete","artifacts":["measurements.csv"]}
```

Artifacts must be regular files beneath the staging directory. Absolute paths,
traversal, symbolic links, duplicate aliases and reserved receipt names are
rejected. Input checksums are verified after execution. Successful output gets a
receipt containing the exact request, settings and request hash and is published
to a new destination. Failed or timed-out jobs do not
publish a result directory. The executable is trusted code: the subprocess
protocol is not a security sandbox or a rollback mechanism for worker side effects.

```{r worker, eval=requireNamespace("processx", quietly=TRUE)}
worker <- file.path(work, "worker.R")
writeLines(c(
  'args <- commandArgs(TRUE)',
  'request <- jsonlite::read_json(args[1])',
  'writeLines("synthetic fixture", file.path(args[2], "fixture.txt"))',
  'jsonlite::write_json(list(schema_version=1, status="complete",',
  '  artifacts=list("fixture.txt")), file.path(args[2], "result.json"),',
  '  auto_unbox=TRUE)'
), worker)
backend <- bs_backend(file.path(R.home("bin"), "Rscript"), args = worker,
                      version = "synthetic-protocol-example-v1")
job <- bs_analyze(c(image = image), backend, out = file.path(work, "job"))
job$status
```

This fixture tests transport and publication only. A scientific worker must also
record model and code identity, calibration, training/test split, valid support,
normalization branch, thresholds, and independent validation. Partition by the
physical source slide before generating training crops to avoid leakage. Sparse
points need an explicit supervision mask; unlabeled pixels are not negatives.

Native QuPath save/export integration, portable DNN training and whole-slide
inference remain separate qualification work. The preserved study scripts and
models are not silently embedded as universal defaults in this package.

```{r cleanup, include=FALSE}
unlink(work, recursive = TRUE)
```
